\documentclass [letterpaper,11pt,twoside]{article}
\input {header-small}

\begin {document}
\setcounter {section}{0}\section {{}, February 4, 2011}
  Geometry of solutions to sets of polynomial equations.

  e.g., $x^2 + y^2 + 1 = 0$ $\longrightarrow$ Set of solutions (over $\mathbb C$) is really a sphere (with two points removed)

  e.g., $y^2 = x^3 - x = x(x-1)(x+1)$
  Over $\mathbb C$, the set of solutions is a torus (with one point removed)

  Lots of applications to number theory, representation theory, etc.

  We'll work over the field $\mathbb C$.

  Recall (background reading, \S 10-7, 10-8 in Artin's \emph{Algebra}):
  \begin{thm*}[Hilbert's Nullstellensatz (weak)]
    The maximal ideals of $\mathbb C[x_1, \ldots, x_n]$ are exactly those of the form $(x_1 - a_1, \ldots, x_n - a_n)$ corresponding to points $(a_1, \ldots, a_n)\in \mathbb C^n$.
  \end{thm*}

  This means we can consider $\mathbb C^n$ as a purely algebraic object.  It's called affine $n$-space, $\mathbb A^n_{\mathbb C}$ or $\mathbb A^n$ for short.

  We want to define a nice topology on this space.  One choice is to take the Euclidean (complex) topology: define open balls by
  $$B_r(x) = \mathset[y\in\mathbb C^n]{|y - x| < r}$$
  and take these to be a basis.

  But this is too many open sets (closed sets), e.g. $\mathset[(x,y)\in\mathbb C^n]{y=e^x}$ is closed in the Euclidean topology.  But we only care about polynomials, so we'll use a coarser topology (fewer open/closed sets).

  We'll use the smallest topology such that polynomial functions are continuous.  This is called the Zariski topology.  Defined by:
  for a polynomial function $f\in \mathbb C[x_1, \ldots, x_n]$, define $D(f) = \mathset[(a_1, \ldots, a_n)\in\mathbb C^n]{f(a_1, \ldots, a_n) \ne 0}$ and declare all $D(f)$ to be open.  (Note: $D$ stands for distinguished.)  As $f$ varies over all polynomials, these $D(f)$ are taken to be a basis.

  We have, e.g, $D(0) = \emptyset$, $D(1) = \mathbb C^n$, $D(fg) = D(f) \cap D(g)$.

  Alternatively, let's see what the closed sets are.  For every ideal $I \subseteq \mathbb C[x_1, \ldots, x_n]$, the vanishing locus of $I$ is $V(I) = \mathset[(a_1, \ldots, a_n)\in\mathbb A^n]{f(a_1, \ldots, a_n) = 0\ \forall f\in I}$.  As $I$ varies, these describe all the closed sets in the Zariski topology.

  Note that $\mathbb C^n \setminus D(f) = V(f)$.

  \textsc{Note}:
  \begin{enumerate}
    \item Because $\mathbb C[x_1, \ldots, x_n]$ is noetherian\footnote{Every ascending chain of ideals stabilizes: given $I_1 \subseteq I_2 \subseteq \cdots \subseteq R$, $\exists k$ such that $I_k = I_{k+ 1} = \cdots$.  Equivalent to that every ideal is finitely generated}, any $I$ can be written as $(f_1, \ldots, f_k)$ for some $f_1, \ldots, f_k$.  So $V(I) = V(\{f_1, \ldots, f_k\})$.
    \item The maximal ideals (the smallest non-empty closed sets) exactly correspond to the points of $\mathbb C^n$.  (weak Nullstellensatz)
  \end{enumerate}

  e.g.
  \begin{enumerate}[1)]
    \item $\mathbb A^1$: the closed sets are $\emptyset = V(1)$, $\mathbb A^1 = V(0)$, and sets of zeros of polynomials, that is, all finite sets of points.  (also called the cofinite topology)
    \item $\mathbb A^2$: the closed sets are $\emptyset = V(1)$, $\mathbb A^1 = V(0)$, finite sets of points, but also union of $V(f)$ with a finite point set for some polynomial $f\in \mathbb C[x,y]$.

      Why don't we include $V(f, g)$ for $f, g \in \mathbb C[x,y]$?  It's because the locus of points where both $f$ and $g$ vanish (assuming they have no common factor) is a finite set of points.  Note also that $V(f_1) \cup V(f_2) = V(f_1 f_2)$.
  \end{enumerate}
  $\mathbb C[x_1, \ldots, x_n]$ is called the affine coordinate ring of $\mathbb A^n$.  (think of it as set of functions on the space $\mathbb A^n$)

  \subsection{Projective Space}
  Claim: $\mathbb A^2$ is somewhat defective.
