\documentclass [letterpaper,11pt,twoside]{article}
\input {header-small}
\usepackage[margin=1in]{geometry}

\usepackage{cancel}
% email to abhinav@math.mit.edu
\def\d{\, {\rm d}}

%\usepackage{pgf,tikz}
%\usetikzlibrary{arrows}
\usepackage{wasysym}
\usepackage{pdfcomment}

\usepackage{datetime}
\usepackage{verbatim}
\usepackage[all]{xy}
\def\classnumber{18.721}
\def\classname{Algebraic Geometry}
\edef\cheadcontent{\classnumber\space Notes}

\pagestyle{fancy}
\headheight 13.7pt
\fancyhead{}
\fancyhead{Jason Gross}%\today}
\fancyfoot{}
\lhead{Jason Gross}
\rhead{\TeX ed on \today}
\chead{\cheadcontent}
\cfoot{\thepage\space of \pageref{LastPage}}

\newcommand{\flag}[2][]{#2\footnote{#1}}

\begin {document}
\setcounter {section}{15}\section {Friday, March 11, 2011}
  Finite group action on an integrally closed finite-type domain $B$, $A = B^G$, (has the?) invariants
  \begin{itemize}
    \item[] $\max A \leftrightarrow G$-orbit in $\max B$
    \item[] $A$ is finite type and integrally closed
    \item[] $B$ is a finite $A$-module
  \end{itemize}
  \begin{defn*}
    $A$ is \emph{normal} if it's integrally closed in the field of fractions of $A$ (i.e., it's own fraction field).
  \end{defn*}
  \begin{thm*}
    Let $G$ be a finite group that acts on an integrally closed finite-type domain $B$, $A = B^G$.  Then
    \begin{itemize}
      \item $\pspec A \leftrightarrow (\pspec B) / G$
      \item This preserves inclusions.  More formally, If $P\leftrightarrow \text{orbit }\{Q_j\}$ and $P' \leftrightarrow \text{orbit }\{Q'_i\}$, then $P \subset P'$ if and only if $\forall j \exists i$ such that $Q_j \subset Q'_i$. \par\noindent
      We have $Q \leadsto A \cap Q$, $\spec B \to \spec A$
    \end{itemize}
  \end{thm*}
  \begin{lem*}
    Let $Q_1, \ldots, Q_m$; $Q_1', \ldots, Q_n'$ be prime ideals of $B$.  Suppose $Q_j \not\subset \text{any }Q_i'$, $j = 1, \ldots, m$, $i = 1, \ldots, n$.  Then there exists $\alpha \in B$, $\alpha \in Q_1, \ldots, Q_m$, $\alpha \notin Q_1', \ldots, Q_n'$.
  \end{lem*}
  \begin{proof}
    Plan: Solve for a single $Q_j$.  Find $\alpha_j\in Q_j$, $\notin Q_1', \ldots, Q_n'$.  Then the product $(\alpha_1 \cdots \alpha_m$) works; $\alpha_1 \cdots \alpha_m \in Q_i' \implies $ some $\alpha_j \in Q_i'$.  This would be a contradiction.

    Take $Q_1 = Q$, $Q_1', \ldots, Q_n'$, $\alpha \in Q$, $\notin Q_i'$.

    %Hypothesis: $Q \not\subset Q_i'$; Conclusion: $Q \not\subset \bigcap Q_i'$.

    $\alpha = \beta_i\in Q$, $\notin Q_i'$, $i = 1,\ldots$. $\beta_1 \cdots \beta_n\in Q$.
    \ldots (Proof not included)
  \end{proof}
  \begin{proof} (of theorem) \par\noindent
    In $C$:
    $$
    \xymatrix @R=1em{
      B \ar[r]^\sigma & B \\
      A \ar@{^{(}->}[u] \ar[r]^{\text{id}} &  A
    }
    $$

    Since $Q$ is a prime ideal of $B$ and $P = A \cap Q$, we have that $\sigma Q$ is a prime ideal of $B$ and $P = A\cap \sigma Q$.  Then an orbit of $Q$ corresponds to one point in $\pspec A$.  By the lying over theorem, $\spec B \to \spec A$ is surjective.  Therefore $(\spec B) / G \to \spec A$ is also surjective.

    \emph{Injective}:   $\{Q_1, \ldots, Q_m\}$, $\{Q_1', \ldots, Q_n'\}$ distinct orbits.  $A \cap Q_j = P$, $A \cap Q_i' = P'$, show $P \ne P'$.
    \begin{claim*}
      We can't have $Q_i \subset Q_j'$ for some $i$, $j$ and also $Q_k' \subset Q_\ell$ for some $k$, $\ell$.
    \end{claim*}
    \begin{proof}
      We can renumber, $i = j = 1$.  $Q_k' \subset Q_\ell \implies \sigma Q_k' \subset \sigma Q_\ell$.  We may then assume that $\ell = 1$; since $\sigma$ runs through the whole orbit, we can choose an appropriate $\sigma$.

      Then $Q_1 \subset Q_1'$, $Q_k' \subset Q_1$.  Then $Q_k' \subset Q_1 \subset Q_1'$.  Then $Q_k' \subset Q_1'$.  Thus $k = 1$, because $Q_k' = \sigma Q_1'$ for some $\sigma$ (and permutations can't take sets to proper subsets of themselves).  Then $Q_1' \subset Q_1 \subset Q_1'$, so $Q_1 = Q_1'$.  This is a contradiction, since orbits are disjoint.  Thus, we may suppose that $Q_j \not\subset Q_i'$ for any $i$, $j$.
    \end{proof}
    There exists an $\alpha \in Q_1 \cap \cdots \cap Q_m$, $\notin Q_i'$.  Take $\gamma = \prod_\sigma \sigma \alpha $, $\gamma\in Q_j$, $\notin Q_i'$, and $\gamma \in A$.  Thus, $\gamma \in P$, $\notin P'$, so $P \ne P'$.

    The proof of the second point is skipped  (See notes that Prof. Artin posted).
  \end{proof}

  Let $B$ be a finite $A$-module for a normal $A$.
    $$
    \xymatrix @R=1em{
      A \ar@{^{(}->}[r] & B \\
      P' \ar@{^{(}->}[u] &  Q' \ar@{^{(}->}[u] \\
      P \ar@{^{(}->}[u] & Q \ar@{^{(}->}[u]
    }
    $$
  $P' = A \cap Q'$, $P = A \cap Q$

  \begin{thm*}[Going Up]
    In the diagram above, given $P$, $P'$, $Q$, there exists a $Q'$.
  \end{thm*}

  \begin{thm*}[Going Down]
    In the diagram above, given $P$, $P'$, $Q'$, there exists a $Q$.
  \end{thm*}
  \begin{proof}
    Let $K$ be the fraction field of $A$ and $L$ be the fraction field of $B$.
    \begin{enumerate}[{Case }1:]
      \item $L / K$ is Galois, has Galois group $G$(?).  $B$ is normal. \par\noindent
        $\sigma$ acts on $B$: The elements of $B$ are integral of $A$.  If $\beta$ is integral, so is $\sigma \beta$.  Then, since $B$ is normal, $B$ is the integral closure of $A$ in $L$.  Thus, $\sigma \beta \in B$.

        $A = B^G$. (We know that $B$ is a finite $B^G$-module.):  $A \subset B^G$ and $A$ is normal.  Since $B$ is integral over $A$, $B^G$ is integral over $A$, so $A = B^G$.

        Then the theorem follows from the previous theorem.
      \item (general case): $L / G$ not Galois, and/or $B$ not normal:  Put $L \subset F$ a Galois extension of $K$ with Galois group $G$.  Let $C$ be the integral closure of $A$ in $F$.  This is a finite $A$-module.  Then $G$ operates on $C$.  Then $A = C^G$ (by the same reasoning as in case 1).

      $$
      \xymatrix @R=0.25em{
        A \ar@{^{(}->}[r] & B \ar@{^{(}->}[r] & C \\
        P'  &  Q' & R'\\
        P  &  & R
      }
      $$

      By lying over, there exists a prime ideal $R'$, $B \cap R' = Q'$.  Case 1 says that there exists an $R$.  Then $A \cap R = P$.  Put $Q = B \cap R$.
    \end{enumerate}
  \end{proof}
\end {document}

