\documentclass [letterpaper,11pt,twoside]{article}
\input {header-small}
\usepackage[margin=1in]{geometry}

\usepackage{cancel}
% email to abhinav@math.mit.edu
\def\d{\, {\rm d}}

%\usepackage{pgf,tikz}
%\usetikzlibrary{arrows}
\usepackage{wasysym}
\usepackage{pdfcomment}

\usepackage{datetime}
\usepackage{verbatim}
\usepackage[all]{xy}
\def\classnumber{18.721}
\def\classname{Algebraic Geometry}
\edef\cheadcontent{\classnumber\space Notes}

\pagestyle{fancy}
\headheight 13.7pt
\fancyhead{}
\fancyhead{Jason Gross}%\today}
\fancyfoot{}
\lhead{Jason Gross}
\rhead{\TeX ed on \today}
\chead{\cheadcontent}
\cfoot{\thepage\space of \pageref{LastPage}}

\newcommand{\flag}[2][]{#2\footnote{#1}}

\begin {document}
\setcounter {section}{18}\section {Friday, March 18, 2011}
  $GK$-dimension $A$ (finite-type algebra, domain) \\
  Choose $V$ a one-dimensional subspace which
  \begin{itemize}
    \item generates $A$
    \item contains 1 (that is, $1\in V$)
  \end{itemize}
  $V^n$ is the span of products of length $n$ of elements of $V$. \\
  $V \subset V^2 \subset \cdots$ \\
  $\bigcup V^n = A$ \\
  The generating function $g$ is $g(n) = \dim V^n$ \\
  $GK$-$\dim V = d$ means that $g(n) \le c n^d$ and $\not\le c n^{d-\epsilon}$

  \begin{lem*} \par\noindent
    \begin{itemize}
      \item This is independent of the choice of $V$.
      \item $\gk(k[x_1, \ldots, x_n]) = d$ (?)
    \end{itemize}
  \end{lem*}
  Take $V = $ polynomials of degree $\le 1$.  Then $\dim V^n = \binom{n+d}{d}$.

  Trivial fact: If $\bar A = A / I$ then $\gk(\bar A) \le \gk(A)$.
  \begin{lem*}
    Suppose $A \subset B$, $B$ a finite $A$-module.  Then $\gk(B) = \gk(A)$.
  \end{lem*}
  \begin{proof}
    Choose a finite dimensional subspace $U$ of $B$ that generates $B$ as an $A$-module.  Then $U^2 \subset B = AU$.  Now choose a (one-dimensional) generating space $V$ for $A$ large enough so that $U^2 \subset VU$.  Take $W = V + U$ to generate $B$ as an algebra.  Then we have
    $$W^n = V^n + V^{n-1} U + V^{n-2}U^2 + \cdots + U^n.$$
    Since $V^{n-2}U^2 \subset V^{n-1}U$, by induction, $U^n \subset V^{n-1}U$.  Then $W^n \subset V^n + V^{n-1}U$.  Since $W = V + U$, we also have that $V^n \subset W^n$.

    Let $h(n) = \dim W^n$.  Then
    $$g(n) \le h(n) \le g(n) + rg(n-1) \le (r+1)g(n)$$
    with $r = \dim U$.
  \end{proof}

  Let $K$ be the fraction field of $A$.
  Define $\trdeg A = \trdeg K / \mathbb C$.\footnote{$\trdeg$ is the transcendence degree.}

  \begin{cor*}
    $\gk(A) = \trdeg K / \mathbb C$.
  \end{cor*}
  \begin{proof}
    Noether Normalization says that $A$ is a finite module over $k[x_1, \ldots, x_k]$, $k[x] \subset A$.  Then $k = \trdeg K / \mathbb C$.
  \end{proof}
  Corollaries:
  \begin{itemize}
    \item $\gk(A) \in \mathbb N$
    \item $s \ne 0, s\in A \implies \gk(A_s) = \gk(A)$ (same field of fractions, therefore same degree)
  \end{itemize}


  Let $A$ be a finite type domain, $P$ a prime ideal of codimension 1 ($(0) < P$ and no primes in between).

  \begin{prop*}
    $\gk(A / P) = \gk(A) - 1$ or $\trdeg A / P = \trdeg A - 1$
  \end{prop*}
  \begin{proof}
    Replace $A$ by its normalization $A'$, a finite $A$-module.


    $$
    \xymatrix @R=0.25em{
      A \ar@{^{(}->}[r] & B \\
      P  &  Q\quad\text{also codim 1 ($P = A \cap Q$)}\\
      (0)  & (0)
    }
    $$
    $$\trdeg A = \trdeg B$$
    $B / Q$ is a finite $A / P$-module, $\trdeg B / Q = \trdeg A / P$.

    Assume $A$ is normal.  Then the local ring $A_P$ is a DVR (discrete valuation ring).  The maximal ideal $P_P = tA_P$ for some $t\in A_P$.  It's a principal ideal.  $t$ is a fraction $s^{-1}a$, $s\notin P$.

    Localize: Replace $A$ by $A_s$, $P$ by $P_s$.  (It still has codimension 1 because the prime ideals in the localized ring are subsets of the original prime ideals.(?))  Now $t \in A$, but it might not generate $P$.  Say $P = (u_1, \ldots, u_r)$.  In the local ring, $t\mid u_i$.  Then, we can get a common denominator, so $u_i = t(s_1^{-1}b_i)$, $b_i \in A$, $s\notin P$.  Now replace $A$ by $A_{s_1}$.  Now $t$ generates $P$.

    We've reduced to the case that $P = tA$ ($A$ normal).  Extend $t$ to a transcendence basis\footnote{\emph{Transcendence basis}: Maximally algebraically independent set in the fraction field.} for the fraction field of $A$, $(t, x_1, \ldots,x_k)$, $x_i \in A$.

    Now $k[t, x_1, \ldots, x_k] \subset A$ and the elements of $A$ are algebraic over $k[t, x]$.  Look at $k[x_1, \ldots, x_k] \to \bar A = A / P = A / tA$.  What's the kernel?  $f(x) \leadsto 0$ means that $f(x) = t\alpha$, $\alpha \in A$.

    \ldots We seem to not have a proof.  (To be posted online.  We'll assume the proposition is true.)
  \end{proof}

  \begin{thm*}
    $\gk(A) = \trdeg A = \text{Krull}\,\dim A$.
    More precisely, every maximal chin of prime ideals $(0) < P_1 < P_2 < \cdots < P_d$ has length $d = \gk(A) = \trdeg A$.
  \end{thm*}
  \begin{proof}
    We have already done $\gk(A) = \trdeg A$.

    We induct on $d$.  We prove the statement: If a maximal chain has length $d$, then $d = \gk(A)$.  Look at $\bar A = A / P_1$.  A maximal chain in $\bar A$ is $(0) = \bar P_1 < \bar P_2 < \cdots < \bar P_d$ (by the correspondence theorem) has length $d - 1$.  Then the Krull dimension of $\bar A$ is $\gk(A) = d - 1$.  Therefore, $\gk(A) = d$.

    (Where is Krull's Theorem?!)
  \end{proof}
\end {document}

