\documentclass [letterpaper,11pt,twoside]{article}
\input {header-small}
\usepackage[margin=1in]{geometry}

\usepackage{cancel}
% email to abhinav@math.mit.edu
\def\d{\, {\rm d}}

%\usepackage{pgf,tikz}
%\usetikzlibrary{arrows}
\usepackage{wasysym}
\usepackage{pdfcomment}

\usepackage{datetime}
\usepackage{verbatim}
\usepackage[all]{xy}
\def\classnumber{18.721}
\def\classname{Algebraic Geometry}
\edef\cheadcontent{\classnumber\space Notes}

\pagestyle{fancy}
\headheight 13.7pt
\fancyhead{}
\fancyhead{Jason Gross}%\today}
\fancyfoot{}
\lhead{Jason Gross}
\rhead{\TeX ed on \today}
\chead{\cheadcontent}
\cfoot{\thepage\space of \pageref{LastPage}}

\newcommand{\flag}[2][]{#2\footnote{#1}}

\begin {document}
\setcounter {section}{23}\section {Friday, April 8, 2011}
  \subsection*{Category Theory}
    \begin{defn*}
      A \emph{category} is a collection of objects, together with morphisms
      (maps) between objects satisfying the following axioms:
      \begin{itemize}
        \item Composition of morphisms is defined.
          $$
          \xymatrix{
            x \ar[r]^f \ar@/_/[rr]_{f\circ g\text{\makebox[0pt][l]{ is defined}}} & y \ar[r]^g & z
          }
          $$
        \item Associative law: If we have
          $$f \stackrel{f}{\to} y \stackrel{g}{\to} z \stackrel{h}{\to} w,$$
          then $(h \circ g) \circ f = h \circ (g \circ f)$
        \item For all objects $y$, there exists an identity $y
        \xrightarrow{i_y} y$ such that $g\circ i_y = g$, $i_y \circ f = f$ for
        all $f$ and $g$ with the appropriate domains and ranges.
      \end{itemize}
    \end{defn*}

    Examples of categories include the category of: sets; topological spaces, algebras.

    \begin{defn*}
      Let $C$ and $D$ be categories.  Then a \emph{functor} $F$ is a map $C
      \stackrel{F}{\to} D$ which maps $\obj(C) \to \obj(D)$ and maps $\morph(C)
      \to \morph(D)$ and which (1) is compatible with composition and (2) maps
      identity morphisms to identity morphisms.
    \end{defn*}

    \begin{example*}
      (top) $\to$ (sets), $x \leadsto $ underlying set
    \end{example*}


    \begin{example*}
      (groups) $\to$ (abelian groups), $G \to G / (aba^{-1}b^{-1} = 1)$
    \end{example*}

    \begin{example*}
      (top) $\xrightarrow{1-1_q} $ abelian groups, $x \leadsto H_q(x)$
      homology
    \end{example*}


    \begin{example*}
      pointed path connected topologies $\to $ groups, $x\leadsto \pi_1(x)$
      (homology group)
    \end{example*}

    \begin{defn*}
      The \emph{dual category} $C^\circ$ of a category $C$ is defined by:
      $$\obj(C^\circ) = \obj(C)$$
      \begin{eqnarray*}
        \text{A morphism} & & \text{A morphism} \\
        x^\circ \xleftarrow{f^\circ} y^\circ & = & x \stackrel{f}{\to} y \\
        \text{in }C^\circ & & \text{in }C
      \end{eqnarray*}
    \end{defn*}

    \begin{defn*}
      A \emph{contravariant functor} is a functor $F$ on $C$ from $C^\circ \to
      D$.
    \end{defn*}


    \begin{example*}
      (vect) $=$ category of complex vector spaces.  $(\text{vect})^\circ \to
      (\text{vect})$, $v\leadsto v^*$.  This functor takes a vector space to its
      dual space.
    \end{example*}


    \begin{example*}
      (top) $\to$ (abelian groups), $x\leadsto H^q(x)$ cohomology
    \end{example*}



    \begin{example*}
      Let $X$ be a topological space.  Then let (opens) denote the category where the
      objects are open sets in $X$, and the morphisms are inclusions $U \subset
      V$; if $U \subset V$, $\exists! U \to V$, and if $U\not\subset V$,
      $\not\exists U \to V$.
    \end{example*}

    \begin{defn*}
      The \emph{sheaf of functions} $F$ on $X$ is the contravariant functor on
      (opens): $(\text{opens})^\circ \stackrel{F}{\to} (\text{algebras})$, $U
      \leadsto F(U) = \text{complex valued functions with domain $U$}$ (an
      algebra).

      If $V \to U$ (i.e., $V\subset U$), then we can restrict a function on $U$
      to $V$; we get $F(U) \stackrel{\text{rest}_V}{\leftarrow} F(U)$, $f|_V
      \leadsfrom f$.

      \emph{Sheaf Axiom for $F$}: Functions on $U$ can be defined ``locally.''
      More formally, suppose $U$ is open in $X$ and $U^i$ are open subsets of
      $U$ that together cover $U$.  Let $U^{ij}$ denote $U^i \cap U^j$.
      $$U \leftarrow \{U^i\} \leftleftarrows \{U^{ij}\}$$
      $$\xymatrix @R=0.5em{
        U^i \\
        & U^{ij} \ar@{_{(}->}[lu] \ar@{_{(}->}[ld] \\
        U^j
      }$$
      Then, given functions $f^i$ on $U^i$ such that the
      restriction of $f^i$ and $f^j$ to $U^{ij}$ are equal for all $i$, $j$, the
      sheaf axiom requires that there exists a unique function $f$ on $U$ such
      that the restriction of $f$ to $U^i$ is $f^i$.
    \end{defn*}

    \begin{example*}
      Let $X = U$ be the real line.  Let $U^1 = (-\infty, 1)$ and let $U^2 = (0,
      \infty)$.  Then $U = U^1 \cup U^2$.  Let $U^1 \cap U^2 = U^{12}$ ($= (0,
      1)$).  Then if $f^1$ and $f^2$ are functions on $U^1$ and $U^2$ and if the
      restriction of $f^1$ to $U^{12}$ is equal to the restriction of $f^2$ to
      $U^{12}$, then $\exists!$ $f$ on $U$, such that its restriction to $U^i =
      f^i$.
    \end{example*}

    \begin{defn*}
      A \emph{sheaf on a topological space $X$} is a contravariant functor
      $(\text{opens})^\circ \stackrel{M}{\to} C$ (some category), $U \leadsto
      M(U)$ which satisfies the sheaf axiom:

      Suppose $\{U^i\}$ covers $U$ and let $U^{ij} = U^i \cap U^j$.  Given an
      element $\alpha^i \in M(U^i)$, if the restriction\footnote{If $V \to U$,
      then $M(V) \xleftarrow{\text{``restriction''}} M(U)$.} of
      $\alpha^i$ and $\alpha^j$ to $M(U^{ij})$ are equal for all $i$, $j$, then
      $\exists!  \alpha \in M(U)$ whose restriction to $U^i$ is $\alpha^i$ for
      all $i$.
    \end{defn*}

    We want to (eventually) write the sheaf axiom more compactly.  First, we
    rewrite it in terms of $M(U)$.  Recall
    $$\xymatrix @R=0.5em{
      U^i \\
      & U^{ij} \ar@{_{(}->}[lu] \ar@{_{(}->}[ld] \\
      U^j
    }$$
    Then
    $$0 \to M(U) \to \prod_i M(U^i) \stackrel[d_1^*]{d_0^*}{\rightrightarrows} \prod_{i,j} M(U^{ij})$$
    $$U \leftarrow \{U^i\} \stackrel[d_1]{d_0}{\leftleftarrows} \{U^{ij}\}$$
    Then the sheaf axiom says that the above sequence is exact\footnote{A
    sequence is exact if the kernel of each map is equal to the image of the
    preceding map.} if we replace $\rightrightarrows$ by the difference $d_0^*
    - d_1^*$.\footnote{We're assuming that the category we're mapping into has a
    $-$ (e.g., that of abelian groups), i.e., that each $M(U)$ has a zero
    element and subtraction.}

    Think about for next time: The structure sheaf on $X = \mathbb A^1$ (with the Zariski
    topology): $U$ open $= X - S$ (with $S$ a finite set) together with $U =
    \emptyset$.  If $S = (s_1, \ldots, s_k)$.  Let $f = (x_1 - s_1)\cdots(x -
    s_k)$, $\mathcal O(U) = \mathbb C[x][f^{-1}]$.  Check the sheaf axiom.
\end {document}

