\documentclass [letterpaper,11pt,twoside]{article}
\input {header-small}
\usepackage[margin=1in]{geometry}

\usepackage{cancel}
% email to abhinav@math.mit.edu
\def\d{\, {\rm d}}

%\usepackage{pgf,tikz}
%\usetikzlibrary{arrows}
\usepackage{wasysym}
\usepackage{pdfcomment}

\usepackage{datetime}
\usepackage{verbatim}
\usepackage[all]{xy}
\def\classnumber{18.721}
\def\classname{Algebraic Geometry}
\edef\cheadcontent{\classnumber\space Notes}

\pagestyle{fancy}
\headheight 13.7pt
\fancyhead{}
\fancyhead{Jason Gross}%\today}
\fancyfoot{}
\lhead{Jason Gross}
\rhead{\TeX ed on \today}
\chead{\cheadcontent}
\cfoot{\thepage\space of \pageref{LastPage}}

\begin {document}
\setcounter {section}{9}\section {Friday, February 25, 2011}
   \noindent \emph{Recall}: \\
  If $I$ is an ideal of $A$, then $\operatorname{rad}I = \mathset[x \in A]{x^n \in I\text{, some }n > 0}$ \\
  $V(I) = V(\operatorname{rad}I)$ \\
  $V(I) \supset V(J) \iff I \subset \operatorname{rad}J$

  irreducible closed set $C$: $C \ne C_1 \cup C_2$, $C_i < C$.

  \begin{prop*}
    $I$ an ideal, $V(I)$ irreducible $iff$ $\rad I$ is prime ideal
  \end{prop*}

  \begin{thm*}
    Let $A$ be a finite-type noetherian ring, $I$ an ideal.  Then $\rad I$ is the intersection of a finite number of prime ideals $\rad I = P_1 \cap \cdots \cap P_k$.  Then we can organize $P_i \not\subset P_j$ for $i \ne j$.  Then we have that $V(I) = V(P_1) \cup \cdots \cup V(P_k)$; $V(I)$ is a finite union of irreducible closed sets.
  \end{thm*}
  \noindent $\spec A = \{\text{maximal ideals}\}$ \\
  $V(I) = \{\mathfrak M\text{ that contains }I\}$ \\

  $I$ an ideal.  Then $\bar A = A / I$.
  \begin{align*}
    A & \to \bar A \\
    I & \leadsto (\bar 0) \\
    \rad I & \phantom{\to} \rad(\bar 0) \\
    & = \text{\emph{nilradical}} \\
    & = \mathset[x]{x^n = 0}
  \end{align*}
  The following are equivalent:
  \begin{itemize}
    \item $\rad I = P_1 \cap \cdots \cap P_k$
    \item $\rad(\bar 0) = \bar P_1 \cap \cdots \cap \bar P_k$
    \item[] (in $\bar A$)\quad also prime ideals
  \end{itemize}
  \begin{proof}
    Suppose the theorem is false for some $A$ and some $I$.  Let $S = \{\text{ideals }J\text{ of }A\text{ such that }\rad J$ is not a finite intersection of prime ideals$\}$.  By hypothesis, $S \ne \emptyset$.  Then there exists a maximal element $I$ in $S$.  Then $\rad I$ not a finite intersection of prime ideals, but every larger ideal is a finite intersection of prime ideals.

  \noindent \emph{Note}:
  \begin{itemize}
    \item $I = \rad I$
    \item $I$ is not a prime ideal
  \end{itemize}
  Then there exist ideals $K$, $L$, $K\not\subset I$, $L \not\subset I$, but $KL \subset I$.  Replace $K$ by $K + I$, $L$ by $L + I$.  Then we have
  $$(K + I)(L + I) \subset KL + KI + IL + II \subset I.$$
  (so it's ok to do these replacements).

  Now $K, L > I$.  Replace $K$ and $L$ with their radicals.
  $$(\rad K)(\rad L) \subset \rad(KL) \subset \rad I = I$$
  (so it's ok to these replacements.)

  We have
  \begin{align*}
    \rad K & = P_1 \cap \cdots \cap P_r \\
    \rad L & = Q_1 \cap \cdots \cap Q_s
  \end{align*}
  \ldots

  We made a mistake somewhere.  (Supposed to replace $I$ by a zero ideal?)  The version on the web is probably correct.  Let's skip this for now.

  \end{proof}

  \subsection*{Finite Group Action}
    Let $B$ be a finite type domain, $G$ the finite group of automorphisms of $B$.  Let $A = B^G$.

    \begin{thm*}
      \begin{itemize}
        \item $A$ is a finite type domain
        \item $G$ operates on $\spec B = Y$
        \item $Y$ maps to $\spec A = X$.  $Y \to X$ is surjective, and the fibers are the $G$-orbits in $Y$
      \end{itemize}
    \end{thm*}
    \begin{example*}
      $B = \mathbb C[x, y]$, $\sigma (x) = -x$, $\sigma(y) = -y$.  Then $\langle \sigma \rangle$ is a group of order 2.

      The invariant functions are $u = x^2$, $v = y^2$, and $w = xy$.  It's not hard to see that every invariant function can be written as some combination of these.

      $A = B^G = \mathbb C[u, v, w] / (w^2 - uv)$

      Then $\spec A = X = \text{locus of }w^2 = uv$ in $\mathbb A^3_{u,v,w}$.  This is a double cone in 3-space.
    \end{example*}
    \begin{proof}
      \noindent \emph{$A$ finite type domain}: \par
      Take $\beta \in B$, orbit $\{b_1 = \beta, \ldots, b_r\}$.  Let $p(x) = (x - b_1)\cdots (x - b_r) = x^r - s_1(b) x^{r-1} + \cdots \pm s_r(b)$.  We have that $\beta$ is a root of $p(x)$.  Since the $s_i(b)$ are symmetric functions, $s_i(b) \in B^G = A$.

      Since $B$ is a finite type domain, say $B$ is generated as an algebra by $\beta_1, \ldots, \beta_m$.  Then each $\beta_i$ is a root of the polynomial, coefficients in $A$.

      Let $A_0 = \mathbb C$-algebra generated by these roots.  Then $A_0$ is a finite-type domain contained in $B$.  Every element in $B$ is a polynomial in $\beta_1, \ldots, \beta_m$.

      If a polynomial with $\beta_i$ as a root has degree $d_i$, then we only need monomials in $\beta_i$ with degree $\le d_i$.  There are only a finite number of monomials in $\beta_i$ of degree $\le d_i$.  Then $B$ is spanned as an $A_0$-module by these monomials.  Thus, $B$ is a finite $A_0$-module.

      $A_0 \subset A \subset B$.  Since $A_0$ is a finite-type domain, $A_0$ is noetherian.  $B$ is a finite-type $A_0$-module.  $A$ is a submodule.  Therefore, $A$ is a finite-type $A_0$ module.  Thus, $A$ is a finite-type algebra.
    \end{proof}
\end {document}

